---
title: "StrathE2E2 Project"
author: "Avinash Pothuri"
date: "  "
output: html_document
---

<br><br>

## Overview 

### Aim

The Harvest Ratio is the proportion of average biomass captured per unit time. It is calculated by both Activity and Power of the fishing gear.

The initial Goal of the project was to create a program that calculates the required Activity multiplier for a specific input target harvest ratio.

This is important because ___

<br><br>

### Report Main

#### Background


For each fishing guild, the harvest ratio is given by aggregating the contribution of all individual fishing gears. This relationship for the first guilds harvest ratio is shown in Equation 1:

$$
A_1 a_1 + B_1 a_2 + \dots + L_1 a_{12} = HR_1
\tag{1}
$$

where:

- \(A_1, B_1, \dots, L_1\) represent the contribution from each fishing gear.
- \(a_1, a_2, \dots, a_{12}\) are the activity multipliers for fishing gears 1 to 12, respectively.
- \(HR_1\) is the total harvest ratio for the first fishing guild.

A similar equation can be found for all 10 fishing guilds, and thus altogether can be written as a matrix equation shown in Equation 2:

$$ 
\mathbf{M} \ \mathbf{x} \ = \ \mathbf{y}
\tag{2}
$$


Where **M** represents the contribution matrix, with each row corresponding to a fishing guild and each column corresponding to an individual fishing gear. The vector **x** contains the activity multipliers for the fishing gears, while **y** contains the resulting harvest ratios for each fishing guild.

$$
\mathbf{M} =
\begin{bmatrix}
A_1 & B_1 & . & . & . & L_1 \\
A_2 & B_2 & . & . & . & L_2 \\
. & . & . &  &  & . \\
. & . &  & . &  & . \\
. & . &  &  & . & . \\
A_{10} & B_{10} & . & . & . & L_{10} \\
\end{bmatrix}
\qquad
\mathbf{x} =
\begin{bmatrix}
a_1 \\
a_2 \\
. \\
. \\
a_{12}
\end{bmatrix}
\qquad
\mathbf{y} =
\begin{bmatrix}
HR_1 \\
HR_2 \\
. \\
. \\
HR_{10}
\end{bmatrix}
$$

The program attempts to find the activity multipliers, by setting vector **y** to the target harvest ratios, and then solving equation 2. Generally since there are only 10 equations compared to 12 variables, this equation should be underconstrained and thus an infinite number of solutions, but since activity multipliers have to be greater than zero there might not always be solutions.


<br><br>

#### Program ####

The following code block loads in all relavent libraries and gets the model from StrathE2E2

```{r setup, message = FALSE}

source("functions.R")
library(nnls)
library(bvls)

library(StrathE2E2)
model <- e2e_read("North_Sea","2003-2013")
# results <- e2e_run(model,nyears=3)

Harvest_Ratio <- F_calc_func_2(model)
Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

# Activity <- model$data$fleet.model$gear_activity
Power <- model$data$fleet.model$gear_group_rel_power
# Activity_Multiplier <- model$data$fleet.model$gear_mult

```

<br><br>

The following code contains functions that generate matrix **M**. This is achieved by setting all activity multipliers a2, a3 ... a12 to zero while maintaining a1 = 1, and then calculates the resultant harvest ratio values. These calculated value correspond to A1, A2, A3 ... A10. The process is repeated for each activity multiplier in turn, allowing for matrix **M** to be constructed.

Additionally, the code checks for guilds which '' have all zero values and thus cannot have a hr value, and then removes the relavent row from the matrix and from the target harvest ratio. ''

<br><br>

```{r}


find_HR_matrix_components <- function(target_activity_index, temperary_model) {
  
  number_of_gears = 12
  
  temperary_activity <- numeric(number_of_gears)
  temperary_activity[target_activity_index] <- 1
  temperary_model$data$fleet.model$gear_mult <- temperary_activity
  
  F_calc_func_2(temperary_model)
}

reduce_hr_matrix <- function(hr_matrix, deleted_rows_index) {
  hr_matrix[-deleted_rows_index, ]
}

reduce_target_hr <- function(target_hr, deleted_rows_index) {
  target_hr[-deleted_rows_index]
}

find_HR_matrix <- function(temperary_model) {
  
  number_of_gears = 12
  
  hr_matrix <- sapply(1:number_of_gears, function(gear_id) {
    hr_components <- find_HR_matrix_components(gear_id, temperary_model)
  })
  
  zero_hr_guilds <- which(rowSums(hr_matrix != 0) == 0)
  
  list(
    hr_matrix = hr_matrix,
    zero_hr_guilds = zero_hr_guilds
  )
}

initialize_data <- function(model, Target_Harvest_Ratio) {

  matrix_data <- find_HR_matrix(model)
  
  list(
    Target_Harvest_Ratio = reduce_target_hr(Target_Harvest_Ratio, matrix_data$zero_hr_guilds),
    Equations_Matrix = reduce_hr_matrix(matrix_data$hr_matrix, matrix_data$zero_hr_guilds),
    zero_hr_guilds = matrix_data$zero_hr_guilds
  )
}

```

<br><br>

One way to solve the linear equations is to use the Solve function in R, however ____

The way it gets close to the answer is by minimising the least squares,
It also enforces x > 0 condition.

To find the soluion, or to get as close as you can to the solution, the program uses the nnls package (non negative _____)

$$
\sum_{i=1}^{n}(Ax_i-y_i)^2
$$

<br><br>

```{r}

solve_activity_multiplier_nnls <- function(hr_matrix, target_hr) {
  modified_activity_multiplier <- coef(nnls(hr_matrix, target_hr))
}

```

<br><br>

### Sum of Residual Squares vs Cosine Similarity



<br><br>

```{r}

find_modified_hr <- function(temperary_model, modified_activity_multiplier) {
  temperary_model$data$fleet.model$gear_mult <- modified_activity_multiplier
  F_calc_func_2(temperary_model)
}

check_if_exact_solution <- function(final_activity_multiplier, harvest_ratio_matrix, target_harvest_ratio) {
  residual <- target_harvest_ratio - harvest_ratio_matrix %*% final_activity_multiplier
  tolerance = 1e-10
  norm(residual, type = "2") < tolerance
}

residue_sum_of_squares <- function(hr_matrix, modified_activity_multiplier, target_harvest_ratio) {
  residuals <- target_harvest_ratio - hr_matrix %*% modified_activity_multiplier
  sum(residuals^2)
}

```

<br><br>

<br><br>

```{r}
Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

hr_data <- initialize_data(model, Target_Harvest_Ratio)

Equations_Matrix <- hr_data$Equations_Matrix
Target_Harvest_Ratio <- hr_data$Target_Harvest_Ratio
zero_hr_guilds <- hr_data$zero_hr_guilds

print(paste("There are zero harvest ratios at rows:", zero_hr_guilds))

modified_activity_multiplier <- solve_activity_multiplier_nnls(Equations_Matrix, Target_Harvest_Ratio)
print("The modified activity multiplier is:")
print(modified_activity_multiplier)
print(" ")

has_exact_solution <- check_if_exact_solution(modified_activity_multiplier, Equations_Matrix, Target_Harvest_Ratio)

if (has_exact_solution) {
  print("We found an exact solution")
} else {
  RSS <- residue_sum_of_squares(
    Equations_Matrix,
    modified_activity_multiplier,
    Target_Harvest_Ratio
  )
  print(paste("There are NO exact solutions but the closest we got has a residual sum of squares of", RSS))
}

modified_hr <- find_modified_hr(model, modified_activity_multiplier)
print("There was a final harvest ratio value of:")
print(modified_hr)
print(" ")

```

<br><br>

#### Sensitivity to Changes in the Target Harvest Ratio

To investigate how sensitive the optimisation is to changes in the target harvest ratio, each guild was examined individually. The aim was to determine how altering the target harvest ratio for a single guild affects the quality of the best-fitting solution.

First, the original harvest ratios were calculated and used as the baseline target harvest ratios. For each guild in turn, its target harvest ratio was varied across a range of values while the target harvest ratios for all other guilds were kept fixed.

For each modified target harvest ratio, the optimisation was rerun to determine the activity multipliers that most closely achieved the specified target. The resulting solution was then evaluated by calculating the sum of squared residuals (SSR) between the achieved harvest ratios and the original harvest ratios.

The figures below show how the SSR changes as the target harvest ratio for each guild is varied. The verticle red dashed line indicates the original harvest ratio for that guild. Lower SSR values indicate solutions that remain close to the original harvest ratios, while larger values indicate that the modified target harvest ratio is increasingly difficult to satisfy without deviating from the original solution.

<!-- <br><br> -->

<!-- echo=FALSE, -->

```{r, fig.width=18, fig.height=12}


change_one_target <- function(model, Equations_Matrix, v) {
  
  modified_activity_multiplier <- solve_activity_multiplier_nnls(Equations_Matrix, v)
  SSR <- residue_sum_of_squares(
    hr_data$Equations_Matrix,
    modified_activity_multiplier,
    hr_data$Target_Harvest_Ratio
  )
}

grouplabels <- c(
    "Planktivorous fish",
    "Demersal fish",
    "Migratory fish",
    "Benthos_susp-dep",
    "Benthos_carn-scav",
    "Zooplankton_carn",
    "Birds",
    "Pinnipeds",
    "Cetaceans",
    "Macrophytes"
  )

Harvest_Ratio <- F_calc_func_2(model)

hr_data <- initialize_data(model, Harvest_Ratio)

par(mfrow = c(3, 3))

for (i in 1:9) {
  
  x_values <- seq(0, Harvest_Ratio[i]*2, length.out = 1000)
  
  y_values <-sapply(x_values, function(v) {
    hr_data$Target_Harvest_Ratio[i] <- v
    change_one_target(model, hr_data$Equations_Matrix, hr_data$Target_Harvest_Ratio)
  })
  
  plot(
    x_values,
    y_values,
    type = "l",
    main = paste("Sensitivity of", grouplabels[i]),
    xlab = "Target Harvest Ratios",
    ylab = "Sum of Squared Residuals"
  )
  
  grid()
  abline(v = Harvest_Ratio[i], col = "red", lty = 5)
}




```


The results show that at the point where ___.

<br><br>

#### Placeholder Title

```{r}

Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

hr_data <- initialize_data(model, Target_Harvest_Ratio)

rss_values <- numeric(10)
labels <- c("None", 1:9)

for (i in 0:9) {

  if (i == 0) {
    Equations_Matrix <- hr_data$Equations_Matrix
    Target_HR <- hr_data$Target_Harvest_Ratio
  } else {
    Equations_Matrix <- reduce_hr_matrix(hr_data$Equations_Matrix, i)
    Target_HR <- reduce_target_hr(hr_data$Target_Harvest_Ratio, i)
  }
  
  modified_activity_multiplier <- solve_activity_multiplier_nnls(Equations_Matrix, Target_HR)
  
  rss_values[i+1] <- residue_sum_of_squares(Equations_Matrix, modified_activity_multiplier, Target_HR)
}


barplot(rss_values,
        names.arg = labels,
        main = "My Bar Graph",
        xlab = "Category",
        ylab = "RSS Value")



```



<br><br>

### Bounded Least Squared Optimisation

One limitation of the original optimisation approach was that some of the resulting activity multipliers were unrealistically large. To address this issue, the optimisation procedure were modified by incorporating the bvls package, which allows upper bounds to be specified for individual activity multipliers. This constrains the optimisation solution to be a more realistic range and prevents excessively large multiplier values from influencing the results.

The following code defines the optimisation function using the bvls package. The function applies the bounded least squares approach to find the activity multipliers that minimise the residual error while ensuring that each multiplier remains within its specified upper and lower limits.

``` {r}

upper_bound_solution <- function(hr_matrix, target_harvest_ratio, upper_bound) {
  
  lower_bound <- rep(0, 12)
  
  fit <- bvls(
    hr_matrix,
    target_harvest_ratio,
    bl = lower_bound,
    bu = upper_bound
  )
  
  modified_activity_multiplier <- fit$x
}


```

<br><br>


```{r}

Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

hr_data <- initialize_data(model, Target_Harvest_Ratio)
Equations_Matrix <- hr_data$Equations_Matrix
Target_Harvest_Ratio <- hr_data$Target_Harvest_Ratio


x_values <- seq(0, 400, length.out = 2000)

y_values <- sapply(x_values, function(upper_bound) {
  residue_sum_of_squares(
    hr_matrix = Equations_Matrix,
    modified_activity_multiplier = upper_bound_solution(Equations_Matrix, Target_Harvest_Ratio, rep(upper_bound, 12)),
    target_harvest_ratio = Target_Harvest_Ratio
  )
})

log_y <- lapply(y_values, log)

plot(x_values, log_y,
     type = "l",
     xlab = "Upper Bound Values",
     ylab = "Log of Sum of Residue Squares",
     main = "Upper Bound values vs RSS")

abline(h = log(RSS), col = "red", lty = 5, lwd = 2)

grid()

```



```{r}

# calculate_final_rss <- function(model, target_hr) {
#   
#   hr_data <- initialize_data(model, target_hr)
#   
#   modified_activity_multiplier <- solve_activity_multiplier_nnls(
#     hr_data$Equations_Matrix,
#     hr_data$Target_Harvest_Ratio
#   )
#   
#   residue_sum_of_squares(
#     hr_data$Equations_Matrix,
#     modified_activity_multiplier,
#     hr_data$Target_Harvest_Ratio
#   )
# }
# 
# 
# sensitivity_matrix <- matrix(
#   0,
#   nrow = 12,
#   ncol = 9
# )
# 
# Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)
# 
# epsilon <- 1e-6
# 
# base_rss <- calculate_final_rss(model, Target_Harvest_Ratio)
# 
# for (i in seq_len('12')) {
#   for (j in seq_len('9')) {
#     
#     perturbed_model <- model
#     
#     perturbed_model$data$fleet.model$gear_group_rel_power[i, j] <- perturbed_model$data$fleet.model$gear_group_rel_power[i, j] + epsilon
#     
#     new_rss <- calculate_final_rss(
#       perturbed_model,
#       Target_Harvest_Ratio
#     )
#     
#     sensitivity_matrix[i, j] <- (new_rss - base_rss) / epsilon
#     
#   }
# }

```




```{r}



```
