---
title: "StrathE2E2 Project"
author: "Avinash Pothuri"
date: "  "
output: html_document
---

<br><br>

## Overview 

The Harvest Ratio is the proportion of average biomass captured per unit time. It is calculated by both Activity and Power of the fishing gear.

The initial Goal of the project was to create a program that calculates the required Activity multiplier for a specific input target harvest ratio.

<br><br>

### Report

#### Background


For each fishing guild, the harvest ratio is given by aggregating the contribution of all individual fishing gears. This relationship for the first guilds harvest ratio is shown in Equation 1:

$$
A_1 a_1 + B_1 a_2 + \dots + L_1 a_{12} = HR_1
\tag{1}
$$

where:

- \(A_1, B_1, \dots, L_1\) represent the contribution from each fishing gear.
- \(a_1, a_2, \dots, a_{12}\) are the activity multipliers for fishing gears 1 to 12, respectively.
- \(HR_1\) is the total harvest ratio for the first fishing guild.

A similar equation can be found for all 10 fishing guilds, and thus altogether can be written as a matrix equation shown in Equation 2:

$$ 
\mathbf{M} \ \mathbf{x} \ = \ \mathbf{y}
\tag{2}
$$


Where **M** represents the contribution matrix, with each row corresponding to a fishing guild and each column corresponding to an individual fishing gear. The vector **x** contains the activity multipliers for the fishing gears, while **y** contains the resulting harvest ratios for each fishing guild.

$$
\mathbf{M} =
\begin{bmatrix}
A_1 & B_1 & . & . & . & L_1 \\
A_2 & B_2 & . & . & . & L_2 \\
. & . & . &  &  & . \\
. & . &  & . &  & . \\
. & . &  &  & . & . \\
A_{10} & B_{10} & . & . & . & L_{10} \\
\end{bmatrix}
\qquad
\mathbf{x} =
\begin{bmatrix}
a_1 \\
a_2 \\
. \\
. \\
a_{12}
\end{bmatrix}
\qquad
\mathbf{y} =
\begin{bmatrix}
HR_1 \\
HR_2 \\
. \\
. \\
HR_{10}
\end{bmatrix}
$$

The program attempts to find the activity multipliers, by setting vector **y** to the target harvest ratios, and then solving equation 2. Generally since there are only 10 equations compared to 12 variables, this equation should be underconstrained and thus an infinite number of solutions, but since activity multipliers have to be greater than zero there might not always be solutions.


<br><br>

#### Program ####

The following code block loads in all relavent libraries and gets the model from StrathE2E2

```{r setup, message = FALSE}

source("functions.R")
library(nnls)
library(bvls)

library(StrathE2E2)
model <- e2e_read("North_Sea","2003-2013")
# results <- e2e_run(model,nyears=3)

Harvest_Ratio <- F_calc_func_2(model)
Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

# Activity <- model$data$fleet.model$gear_activity
Power <- model$data$fleet.model$gear_group_rel_power
# Activity_Multiplier <- model$data$fleet.model$gear_mult

```

<br><br>

The following code contains functions that generate matrix **M**. This is achieved by setting all activity multipliers a2, a3 ... a12 to zero while maintaining a1 = 1, and then calculates the resultant harvest ratio values. These calculated value correspond to A1, A2, A3 ... A10. The process is repeated for each activity multiplier in turn, allowing for matrix **M** to be constructed.

Additionally, the code checks for guilds which '' have all zero values and thus cannot have a hr value, and then removes the relavent row from the matrix and from the target harvest ratio. ''

<br><br>

```{r}

find_HR_matrix_components <- function(target_activity_index, temperary_model) {
  
  number_of_gears = 12
  
  temperary_activity <- numeric(number_of_gears)
  temperary_activity[target_activity_index] <- 1
  temperary_model$data$fleet.model$gear_mult <- temperary_activity
  
  F_calc_func_2(temperary_model)
}

reduce_hr_matrix <- function(hr_matrix, deleted_rows_index) {
  hr_matrix[-deleted_rows_index, ]
}

reduce_target_hr <- function(target_hr, deleted_rows_index) {
  target_hr[-deleted_rows_index]
}

find_HR_matrix <- function(temperary_model) {
  
  number_of_gears = 12
  
  hr_matrix <- sapply(1:number_of_gears, function(gear_id) {
    hr_components <- find_HR_matrix_components(gear_id, temperary_model)
  })
  
  zero_hr_guilds <- which(rowSums(hr_matrix != 0) == 0)
  
  list(
    hr_matrix = reduce_hr_matrix(hr_matrix, zero_hr_guilds),
    zero_hr_guilds = zero_hr_guilds
  )
}

initialize_data <- function(model, Target_Harvest_Ratio) {

  matrix_data <- find_HR_matrix(model)
  
  list(
    Target_Harvest_Ratio = reduce_target_hr(Target_Harvest_Ratio, matrix_data$zero_hr_guilds),
    Equations_Matrix = matrix_data$hr_matrix,
    zero_hr_guilds = matrix_data$zero_hr_guilds
  )
}

```

<br><br>

The way it gets close to the answer is by minimising the least squares,
It also enforces x > 0 condition.

$$
\sum_{i=1}^{n}(Ax_i-y_i)^2
$$

<br><br>

```{r}

solve_activity_multiplier_nnls <- function(hr_matrix, target_hr) {
  modified_activity_multiplier <- coef(nnls(hr_matrix, target_hr))
}

```

<br><br>

<br><br>

```{r}

find_modified_hr <- function(temperary_model, modified_activity_multiplier) {
  temperary_model$data$fleet.model$gear_mult <- modified_activity_multiplier
  F_calc_func_2(temperary_model)
}

check_if_exact_solution <- function(final_activity_multiplier, harvest_ratio_matrix, target_harvest_ratio) {
  residual <- target_harvest_ratio - harvest_ratio_matrix %*% final_activity_multiplier
  tolerance = 1e-10
  norm(residual, type = "2") < tolerance
}

residue_sum_of_squares <- function(hr_matrix, modified_activity_multiplier, target_harvest_ratio) {
  residuals <- target_harvest_ratio - hr_matrix %*% modified_activity_multiplier
  sum(residuals^2)
}

```

<br><br>

<br><br>

```{r}
Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

hr_data <- initialize_data(model, Target_Harvest_Ratio)

Equations_Matrix <- hr_data$Equations_Matrix
Target_Harvest_Ratio <- hr_data$Target_Harvest_Ratio
zero_hr_guilds = hr_data$zero_hr_guilds

modified_activity_multiplier <- solve_activity_multiplier_nnls(Equations_Matrix, Target_Harvest_Ratio)

has_exact_solution <- check_if_exact_solution(modified_activity_multiplier, Equations_Matrix, Target_Harvest_Ratio)
modified_hr <- find_modified_hr(model, modified_activity_multiplier)

print(modified_activity_multiplier)
```

<br><br>

### Week 2

To find the soluion, or to get as close as you can to the solution, the program uses the nnls package (non negative _____)

<br><br>

``` {r}

upper_bound_solution <- function(hr_matrix, target_harvest_ratio, upper_bound) {
  
  lower_bound <- rep(0, 12)
  
  fit <- bvls(
    hr_matrix,
    target_harvest_ratio,
    bl = lower_bound,
    bu = upper_bound
  )
  
  modified_activity_multiplier <- fit$x
}


```

<br><br>


```{r}

Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

hr_data <- initialize_data(model, Target_Harvest_Ratio)
Equations_Matrix <- hr_data$Equations_Matrix
Target_Harvest_Ratio <- hr_data$Target_Harvest_Ratio


x_values <- seq(1, 3, length.out = 200)

y_values <- sapply(x_values, function(upper_bound) {
  residue_sum_of_squares(
    hr_matrix = Equations_Matrix,
    modified_activity_multiplier = upper_bound_solution(Equations_Matrix, Target_Harvest_Ratio, rep(upper_bound, 12)),
    target_harvest_ratio = Target_Harvest_Ratio
  )
})

plot(x_values, y_values,
     type = "l",
     xlab = "Upper Bound Values",
     ylab = "Sum of Residue Squares",
     main = "Upper Bound values vs RSS")

# abline(h = RSS, col = "red", lty = 5, lwd = 2)

grid()

```




```{r}

Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)

hr_data <- initialize_data(model, Target_Harvest_Ratio)

rss_values <- numeric(10)
labels <- c("None", 1:9)

for (i in 0:9) {

  if (i == 0) {
    Equations_Matrix <- hr_data$Equations_Matrix
    Target_HR <- hr_data$Target_Harvest_Ratio
  } else {
    Equations_Matrix <- reduce_hr_matrix(hr_data$Equations_Matrix, i)
    Target_HR <- reduce_target_hr(hr_data$Target_Harvest_Ratio, i)
  }
  
  modified_activity_multiplier <- solve_activity_multiplier_nnls(Equations_Matrix, Target_HR)
  
  rss_values[i+1] <- residue_sum_of_squares(Equations_Matrix, modified_activity_multiplier, Target_HR)
}


barplot(rss_values,
        names.arg = labels,
        main = "My Bar Graph",
        xlab = "Category",
        ylab = "RSS Value")



```


```{r}

# calculate_final_rss <- function(model, target_hr) {
#   
#   hr_data <- initialize_data(model, target_hr)
#   
#   modified_activity_multiplier <- solve_activity_multiplier_nnls(
#     hr_data$Equations_Matrix,
#     hr_data$Target_Harvest_Ratio
#   )
#   
#   residue_sum_of_squares(
#     hr_data$Equations_Matrix,
#     modified_activity_multiplier,
#     hr_data$Target_Harvest_Ratio
#   )
# }
# 
# 
# sensitivity_matrix <- matrix(
#   0,
#   nrow = 12,
#   ncol = 9
# )
# 
# Target_Harvest_Ratio <- c(0.0001, 0.0004, 0.002, 0.0004, 0.0003, 0.0005, 0.00008, 0.000002, 0.001, 0)
# 
# epsilon <- 1e-6
# 
# base_rss <- calculate_final_rss(model, Target_Harvest_Ratio)
# 
# for (i in seq_len('12')) {
#   for (j in seq_len('9')) {
#     
#     perturbed_model <- model
#     
#     perturbed_model$data$fleet.model$gear_group_rel_power[i, j] <- perturbed_model$data$fleet.model$gear_group_rel_power[i, j] + epsilon
#     
#     new_rss <- calculate_final_rss(
#       perturbed_model,
#       Target_Harvest_Ratio
#     )
#     
#     sensitivity_matrix[i, j] <- (new_rss - base_rss) / epsilon
#     
#   }
# }

```

