---
title: "StrathE2E2 Project"
author: "Avinash Pothuri"
date: "`r Sys.Date()`"
output: html_document
---

# {.tabset}

<br><br>

## Overview

<br><br>

### Aim

This project develops an optimisation framework for determining fishing gear activity multipliers required to achieve specific harvest ratios within the StrathE2E2 ecosystem model. 

The Harvest Ratio is the proportion of average biomass captured per unit time. It is calculated by both Activity and Power of the fishing gear.

<br><br>

## Background / Setup

<br><br>

### Background

For each fishing guild, the harvest ratio is given by aggregating the contribution of all individual fishing gears. This relationship for the first guilds harvest ratio is shown in Equation 1:

$$A_1 a_1 + B_1 a_2 + \dots + L_1 a_{12} = HR_1
\tag{1}$$

where:

- $A_1, B_1, \dots, L_1$ represent the contribution from each fishing gear.
- $a_1, a_2, \dots, a_{12}$ are the activity multipliers for fishing gears 1 to 12, respectively.
- $HR_1$ is the total harvest ratio for the first fishing guild.

A similar equation can be found for all 10 fishing guilds, and thus altogether can be written as a matrix equation shown in Equation 2:

$$ 
\mathbf{M} \ \mathbf{x} \ = \ \mathbf{y}
\tag{2}
$$

Where **M** represents the contribution matrix, with each row corresponding to a fishing guild and each column corresponding to an individual fishing gear. The vector **x** contains the activity multipliers for the fishing gears, while the vector **y** represents the desired harvest ratios for each fishing guild and defines the optimisation target that the activity multipliers must achieve.

$$
\mathbf{M} =
\begin{bmatrix}
A_1 & B_1 & . & . & . & L_1 \\
A_2 & B_2 & . & . & . & L_2 \\
. & . & . &  &  & . \\
. & . &  & . &  & . \\
. & . &  &  & . & . \\
A_{10} & B_{10} & . & . & . & L_{10} \\
\end{bmatrix}
\qquad
\mathbf{x} =
\begin{bmatrix}
a_1 \\
a_2 \\
. \\
. \\
a_{12}
\end{bmatrix}
\qquad
\mathbf{y} =
\begin{bmatrix}
HR_1 \\
HR_2 \\
. \\
. \\
HR_{10}
\end{bmatrix}
$$

The program attempts to find the activity multipliers, by setting vector **y** to the target harvest ratios, and then solving equation 2. Although the linear system is underdetermined (10 equations and 12 unknowns), the activity multipliers must be non-negative, which restricts any feasable solution.

<br><br>

### Setup

the program uses many packages all of which are loaded below

<br><br>

```{r, message = FALSE}

source("functions.R")

library(igraph)
library(lpSolve)
library(nloptr)
library(nnls)
library(bvls)

library(StrathE2E2)

model <- e2e_read("North_Sea","2003-2013")

```

<br><br>

### Generate System of equations

To generate the system of equations we take the power matrix and take into account the current Activity values and Harvest Ratio Scale vector, the following function constructs matrix **M**

<br><br>

```{r}

find_matrix_M <- function(model) {
  
  model_temporary <- model
  
  Power <- as.matrix(model_temporary$data$fleet.model$gear_group_rel_power)
  HRscale <- model_temporary$data$fleet.model$HRscale_vector
  Activity <- model_temporary$data$fleet.model$gear_activity
  
  Power <- Power *Activity
  
  t_Power <- t(Power)
  
  M <- t_Power * HRscale
  
  return(M)
}

```

<br><br>

### Comparing Solutions using Residual Sum of Squares

When the program solves for the activity multipliers there needs to be a way to they if they are good. One way to do that is to find the sum of residue squares of the resultant Harvest Ratio and the target Harvest Ratio.

Additionally, some guilds may have no gear that affects them. In these cases, the value will always be 0 and this should be taken into account.

A function for this is shown below:

<br><br>

```{r}

calculate_rss <- function(model, M, y, gear_multiplier) {
  
  model_temporary <- model
  
  model_temporary$data$fleet.model$gear_mult <- gear_multiplier
  predicted <- F_calc_func_2(model_temporary)

  keep <- rowSums(M != 0) > 0

  predicted <- predicted[keep]
  y <- y[keep]
  
  rss_value <- sum((y - predicted)^2)
  return(rss_value)
}

```

<br><br>

## Graphical Analysis

<br><br>

In this section we explore ways to asnalytically breakdown and solve for the activity multipliers.

<br><br>

### Using graphs to analyse systems of equations

We can decompose this system of simultaneous equations into independent subsystems by representing it as a graph, where the Activity Gears and Harvest Ratios are nodes, and an edge exists between them if and only if the corresponding power value is non-zero. The independent subsystems are then obtained by identifying the connected components of the graph, since each connected component represents a set of equations that is independent of the others.

This special type of graph is called a bipartite graph, in which the vertices are divided into two disjoint sets, and every edge connects a vertex in one set to a vertex in the other. Consequently, there are no edges between vertices within the same set.

For example if we are given the following simultaneous equation

$$
\begin{cases}
2x + 3y = 5 \\
4y = 7
\end{cases}
$$

This can be turned into a matrix equation in the form

$$
\begin{bmatrix}
2 & 3 \\
0 & 4
\end{bmatrix}
\begin{bmatrix}
x \\
y
\end{bmatrix}
=
\begin{bmatrix}
5 \\
7
\end{bmatrix}
$$

Now can turn this into a graph

```{r, echo=FALSE}

vertices <- data.frame(
  name = c("Equation 1", "Equation 2", "x", "y")
)

edges <- data.frame(
  from = c("Equation 1", "Equation 1", "Equation 2"),
  to   = c("x",          "y",          "y")
)

g <- graph_from_data_frame(edges, vertices = vertices, directed = FALSE)

vertex_cols <- ifelse(vertices$name %in% c("x", "y"), "pink", "lightblue")

plot(
  g,
  vertex.size = 35,
  vertex.color = vertex_cols,
  vertex.frame.color = "black",
  vertex.label.cex = 0.9,
  edge.arrow.size = 0.6
)

```


The following code block contains code to build the graph:

<br><br>

```{r}

build_graph <- function(Matrix) {

  hr_names <- paste0("Harvest Ratio", 1:10)
  activity_names <- paste0("Activity Multiplier", 1:12)
  
  nz <- which(Matrix != 0, arr.ind = TRUE)
  
  edges <- data.frame(
    from = hr_names[nz[,1]],
    to   = activity_names[nz[,2]]
  )
  
  g <- graph_from_data_frame(edges, directed = FALSE)
  
  return(g)
}

```

<br><br>

### Finding independent equations

By identifying the connected components of the bipartite graph, we can decompose the original system into independent subsystems of simultaneous equations. Each subsystem can then be solved independently. The following function performs this decomposition and returns a list of coefficient matrices, each representing an independent subsystem of simultaneous equations.

<br><br>

```{r}

find_independent_blocks <- function(g, Matrix) {

  rownames(Matrix) <- paste0("Harvest Ratio", seq_len(nrow(Matrix)))
  colnames(Matrix) <- paste0("Activity Multiplier", seq_len(12))

  comp <- components(g)

  blocks <- lapply(seq_len(comp$no), function(i) {

    nodes <- names(comp$membership[
      comp$membership == i
    ])

    rows <- nodes[grepl("^Harvest Ratio", nodes)]
    cols <- nodes[grepl("^Activity Multiplier", nodes)]

    Matrix[
      match(rows, rownames(Matrix)),
      match(cols, colnames(Matrix)),
      drop = FALSE
    ]
  })

  return(blocks)
}

```

<br><br>

### Removing redundant equations

Now there are three possible outcomes when solving a system of simultaneous equations:

A unique solution.
No solution (the system is inconsistent).
Infinitely many solutions.

Before attempting to solve the system, any redundant equations should be removed. A redundant equation does not provide any additional information because it is a linear combination of the other equations. For example,

$$
\begin{cases}
x + y = 2, \\
2x + 2y = 4.
\end{cases}
$$

The second equation is simply twice the first and therefore does not impose an additional constraint. Removing redundant equations reduces the size of the system without changing its solution set.

The following function identifies and removes redundant equations while preserving all independent constraints. If the original system is inconsistent, the function also preserves the equations required to detect that inconsistency.

<br><br>

``` {r}

remove_redundant_equations <- function(blocks, target_harvest_ratio, tol = 1e-10) {
  
  lapply(blocks, function(M) {
    
    if (nrow(M) <= 1)
      return(M)
    
    X <- as.integer(sub("Harvest Ratio", "", rownames(M)))
    y <- target_harvest_ratio[X]
    
    keep <- integer(0)
    
    for (i in seq_len(nrow(M))) {
      
      candidate <- c(keep, i)
      
      A <- M[candidate, , drop = FALSE]
      b <- y[candidate]
      
      rank_A  <- qr(A, tol = tol)$rank
      rank_Ab <- qr(cbind(A, b), tol = tol)$rank
      
      if (rank_Ab > rank_A || rank_A > length(keep)) {
        keep <- candidate
      }
    }
    
    M[keep, , drop = FALSE]
  })
}

```

<br><br>

## Analytic Methods

<br><br>

After removing redundant equations, the remaining solvable systems can be solved analytically using R's solve() function and the lpSolve package.

<br><br>

### Single Power Terms

If a fishing gear affects only a single guild, and that guild is affected by no other gears, then the corresponding activity multiplier can be found directly by dividing the target harvest ratio by the associated power term.

An example is a power matrix with ones on the main diagonal and zeros elsewhere, combined with a random target harvest ratio, is shown below: 

<br><br>

```{r}

model_temporary <- model

Diagonal <- as.data.frame(diag(1, nrow = 12, ncol = 10))

names(Diagonal) <- names(model_temporary$data$fleet.model$gear_group_rel_power)
rownames(Diagonal) <- rownames(model_temporary$data$fleet.model$gear_group_rel_power)

model_temporary$data$fleet.model$gear_group_rel_power <- Diagonal

Target_Harvest_Ratio <- F_calc_func_2(model) * runif(10, min = 10, max = 100)

M <- find_matrix_M(model_temporary)
G <- build_graph(M)
blocks <- find_independent_blocks(G, M)
blocks <- remove_redundant_equations(blocks, Target_Harvest_Ratio)

Act <- rep(0, 12)

for (i in seq_along(blocks)) {
  if (length(blocks[[i]]) == 1) {
    
    m <- blocks[[i]]

    X <- as.integer(sub("Harvest Ratio", "", rownames(m)))
    Y <- as.integer(sub("Activity Multiplier", "", colnames(m)))

    Act[Y] <- Target_Harvest_Ratio[X] / m[1,1]
  }
}

rss <- calculate_rss(model_temporary, M, Target_Harvest_Ratio, Act)
paste("The rss value is", rss)

```

<br><br>

### Square Power Terms

If the number of equations equals the number of variables and the determinant of the coefficient matrix is nonzero, then the system has a unique solution. Otherwise, there is no solution, since all redundant equations have already been removed.

The unique solution should then be checked to ensure that all values are positive. If any value is non-positive, then no exact feasible solution exists.

The system can be solved using R's built-in solve() function. An example is shown below:

```{r}

# Use better examples, most of these are unsolvable with x > 0

model_temporary <- model

mat <- as.data.frame(matrix(0, nrow = 12, ncol = 10))

mat[1:4, 1:4] <- 1
mat[5:7, 5:7] <- 1
mat[8:9, 8:9] <- 1
mat[10, 10] <- 1

diag(mat) <- 2

names(mat) <- names(model_temporary$data$fleet.model$gear_group_rel_power)
rownames(mat) <- rownames(model_temporary$data$fleet.model$gear_group_rel_power)

model_temporary$data$fleet.model$gear_group_rel_power <- mat

Target_Harvest_Ratio <- c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1)

M <- find_matrix_M(model_temporary)
G <- build_graph(M)
blocks <- find_independent_blocks(G, M)
blocks <- remove_redundant_equations(blocks, Target_Harvest_Ratio)

Act <- rep(0, 12)
solved <- logical(length(blocks))

for (i in seq_along(blocks)) {
  
  m <- blocks[[i]]
  
  X <- as.integer(sub("Harvest Ratio", "", rownames(m)))
  Y <- as.integer(sub("Activity Multiplier", "", colnames(m)))
  
  if (length(m) == 1) {

    Act[Y] <- Target_Harvest_Ratio[X] / m[1, 1]
    
    solved[i] <- TRUE
    
  } else if (nrow(m) == ncol(m)) {
    
    target <-  Target_Harvest_Ratio[X]
    
    if (det(m) != 0) {
      
       sol <- solve(m, target)
       
       if (all(sol >= 0)) {
         
         Act[Y] <- sol
         solved[i] <- TRUE
         
       }
    }
  }
}


unsolved_matrices <- blocks[!solved]

# print(Act)
# print(unsolved_matrices)

```

<br><br>

### More Variables than Equations: Using lpsolve

If there are more variables than equations (i.e. more columns than rows), the system is underdetermined and therefore has infinitely many solutions. However, imposing the additional constraint that all activity multipliers must be positive may eliminate some or all of these solutions.

We can use the lpsolve package which is shown below:

<br><br>

```{r}

solve_nonnegative <- function(M, y) {
  
  m <- nrow(M)
  n <- ncol(M)
  
  obj <- rep(0, n)
  
  sol <- lp(
    direction = "min",
    objective.in = obj,
    const.mat = M,
    const.dir = rep("=", m),
    const.rhs = y
  )
  
  sol
}

#
#
# Add an example of lpsolve using a augmented power matrix
#
#

```

<br><br>

## Computational Methods

<br><br>

### Sum of Residual Squares vs Cosine Similarity

The sum of squared resisiduals (SSR) measures the total squared difference between the predicted and observed values. It is defined as:

$$SSR = \sum_{i=1}^{n}(y_i-\hat{y}_i)^2$$

where $y_i$ are the target values and $\hat{y}_i$ are the corresponding model predictions. A smaller SSR indicates that the predicted values are numerically closer to the observations. Because the residuals are squared, larger errors are penalised more heavily than smaller ones, making SSR particularly sensitive to outliers.

One advantage of SSR is that it is widely used and supported by many optimisation and statistical packages in R.

In contrast, cosine similarity measures the cosine of the angle between two vectors rather than the distance between them. It is defined as:

$$\cos(\theta) = \frac{y \cdot \hat{y}}{|y| |\hat{y}|}$$

where **y** and **$\hat{y}$** are the target and predicted vectors, respectively. Cosine similarity ranges from 0 to 1 for non-negative vectors, with value 1 indicating that the vectors have exactly the same direction.

Cosine Similarity thus doesn't account for the length (magnitude) of the vector, only how closely their directions align.

<br><br>

### Solving System of equations by minimising SSR

One way to solve a system of linear equations in R is to use the **solve()** function. However, **solve()** only works when the the coefficient matrix, **A**, is square, non-singular, and the system has an exact solution. In this case the coefficient matrix is not square and an exact solution does not always exist.

Instead, the solution can be approximated by minimizing the sum of squared residuals. Additionally,the activity multipliers are constrained to be non-negative, as negative values are not physically meaningful.

To find the exact solution (when on exists), or the closest non-negative approximation, the program uses the nnls package in R, which implements the Lawson-Hanson algorithm for No-Negative Least Squares (NNLS). This algorithm finds vector **x** that minimises the sum of squared residuals while satisfying the non-negativity constraint **x** $\geq$ 0:

$$\min_{x \geq 0}\sum_{i=1}^{n}(Ax_i-y_i)^2$$

where **A** is the coefficient matrix, **x** is the vector of activity multipliers, and **y** is the target harvest ratio vector.

<br><br>

```{r}

solve_nnls <- function(M, y) {
  
  solution <- nnls(M, y)
  
  list(
    x = coef(solution),
    SSR = sum(residuals(solution)^2)
  )
}

```

<br><br>

### Solving System of equations by maximising cosine similarity

Cosine Similarity is defined as the cosine of the angle between two vectors, where a value of 1 indicates that the vectors point in the same direction, 0 indicates orthogonality and -1 indicates opposite directions.

For two vectors x and y, cosine similarity in R is implemented below:

<br><br>

```{r}
cosine_similarity <- function(x, y) {
  norm_x <- sqrt(sum(x^2))
  norm_y <- sqrt(sum(y^2))
  
  cos_sim <- sum(x * y) / (norm_x * norm_y)
  return(cos_sim)
}
```

<br><br>

Instead of solving $$Ax=y$$ directly, this program finds x by searching for x where the cosine similarity between **Ax** and **y** is maximised

$$f(x) = \frac{Ax \cdot y}{|Ax| |y|}$$
Since |y| is constant with respect to x, this is equivalent to solving:

$$f(x) = \max_{x \geq 0} \frac{(Ax)^T y}{|Ax|}$$

The optimisation is done using the nloptr library with the NLOPT_LD_SLSQP algorithm, a gradient-descent based constrained optimisation method, which was found to be the most efficient and accurate. The gradient of the objective function is:

$$\nabla_x f(x) = A^T \left(\frac{y}{\|Ax\|} - \frac{((Ax)^T y)Ax}{\|Ax\|^3} \right)$$

The expression is found by using the chain rule to the cosine similarity objective.

<br><br>

```{r}


solve_maximise_cosine_similarity <- function(A, b) {

  cos_obj <- function(x) {
    Ax <- A %*% x
    -sum(Ax * b) / sqrt(sum(Ax^2))
  }

  cos_grad <- function(x) {
    Ax <- A %*% x
    s <- sqrt(sum(Ax^2))

    as.vector(
      t(A) %*% (-b / s + (sum(Ax * b) / s^3) * Ax)
    )
  }

  x <- nloptr(
    x0 = rep(1, ncol(A)),
    eval_f = cos_obj,
    eval_grad_f = cos_grad,
    lb = rep(0, ncol(A)),
    opts = list(
      algorithm = "NLOPT_LD_SLSQP",
      maxeval = 10000,
      xtol_rel = 1e-12
    )
  )
  
  Ax <- A %*% x$solution
  c <- sqrt(sum(b^2)) / sqrt(sum(Ax^2))
  
  c * x$solution
}

```

<br><br>

## Program

<br><br>

### Combining Analytic and Computational methods

The following section combines all previous methods into a single program, which is shown below:

<br><br>

``` {r}

program <- function(blocks, Target_Harvest_Ratio) {
  
  Act <- numeric(12)
  solved <- logical(length(blocks))
  
  for (i in seq_along(blocks)) {
  
    m <- blocks[[i]]
  
    X <- as.integer(sub("Harvest Ratio", "", rownames(m)))
    Y <- as.integer(sub("Activity Multiplier", "", colnames(m)))
  
    if (length(m) == 1) {
  
      V <- m[1, 1]
  
      Act[Y] <- Target_Harvest_Ratio[X] / V
  
      solved[i] <- TRUE
  
    } else if (nrow(m) == ncol(m)) {
  
      target <-  Target_Harvest_Ratio[X]
  
      if (det(m) != 0) {
  
        sol <- solve(m, target)
  
        if (all(sol >= 0)) {
  
          Act[Y] <- sol
          solved[i] <- TRUE
  
        }
      }
    } else if (nrow(m) < ncol(m)) {
  
      targ <- Target_Harvest_Ratio[X]
  
      sol <- solve_nonnegative(m, targ)
  
      if (sol$status == 0) {
        Act[Y] <- sol$solution
        solved[i] <- TRUE
      }
  
    }
  }
  
  unsolved_matrices <- blocks[!solved]
  
  for (m in unsolved_matrices) {
  
    X <- as.integer(sub("Harvest Ratio", "", rownames(m)))
    Y <- as.integer(sub("Activity Multiplier", "", colnames(m)))
  
    target <- Target_Harvest_Ratio[X]
  
    sol <- solve_nnls(m, target)
  
    Act[Y] <- sol$x
  }
  
  return(Act)
}

```

<br><br>

An example includes:

<br><br>

```{r}

model_temporary <- model

Target_Harvest_Ratio <- c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1)

M <- find_matrix_M(model_temporary)
G <- build_graph(M)
blocks <- find_independent_blocks(G, M)
blocks <- remove_redundant_equations(blocks, Target_Harvest_Ratio)

Act <- program(blocks, Target_Harvest_Ratio)

rss <- calculate_rss(model_temporary, M, Target_Harvest_Ratio, Act)
paste("The rss value is", rss)

```

<br><br>

## Power Analysis

<br><br>

### Bridges

The program developed above computes the activity multipliers required to achieve a specified target harvest ratio. If no exact solution exists, it instead finds the best feasible approximation.

Next, we investigate whether removing power values (i.e. setting selected power terms to zero) can transform a previously unsolvable problem into one that admits a solution.

In graph theory, a bridge (also called a cut edge or isthmus) is an edge whose removal increases the number of connected components of a graph. In the context of this problem, removing a bridge separates the system into independent subproblems, which may make them individually solvable.

The following algorithm identifies all relevant bridges in the dependency graph:

<br><br>

```{r}

valid_bridges <- function(g) {

  bridge_edges <- bridges(g)

  if (length(bridge_edges) == 0) {
    return(integer(0))
  }
  
  valid <- integer(0)

  for (e in bridge_edges) {

    g2 <- delete_edges(g, e)

    comp_sizes <- components(g2)$csize

    if (min(comp_sizes) >= 2) {
      valid <- c(valid, e)
    }
  }

  valid
}

```

<br><br>

An example is shown below:

<br><br>

```{r}

model_temporary <- model

mat <- as.data.frame(matrix(0, nrow = 12, ncol = 10))

mat[1:4, 1:4] <- 1
mat[5:7, 5:7] <- 1
mat[8:9, 8:9] <- 1
mat[10, 10] <- 1

diag(mat) <- 2

# The bridges:
mat[5, 4] <- 1
mat[8, 7] <- 1
mat[10, 9] <- 1

names(mat) <- names(model_temporary$data$fleet.model$gear_group_rel_power)
rownames(mat) <- rownames(model_temporary$data$fleet.model$gear_group_rel_power)

model_temporary$data$fleet.model$gear_group_rel_power <- mat

M <- find_matrix_M(model_temporary)
G <- build_graph(M)

valid <- valid_bridges(G)
ends <- ends(G, valid)

print(ends)

```

<br><br>

### Bridges for pairs of edges

Bridges identify single edges whose removal disconnects a graph. However, some graphs can only be disconnected by removing a pair of edges. The following algorithm searches for these edge pairs and is shown below:

<br><br>

```{r}

valid_edge_pairs <- function(g, min_size = 2) {

  edge_ids <- seq_len(ecount(g))

  if (length(edge_ids) < 2) {
    return(list())
  }

  bridge_edges <- bridges(g)

  edge_pairs <- combn(setdiff(edge_ids, bridge_edges), 2, simplify = FALSE)

  valid <- list()

  for (pair in edge_pairs) {

    g2 <- delete_edges(g, pair)

    comp <- components(g2)

    if (comp$no == 2 && min(comp$csize) >= min_size) {
      valid[[length(valid) + 1]] <- pair
    }
  }

  valid
}

```

<br><br>

An example is shown below:

<br><br>

```{r}

model_temporary <- model

mat <- as.data.frame(matrix(0, nrow = 12, ncol = 10))

mat[1:4, 1:4] <- 1
mat[5:7, 5:7] <- 1
mat[8:9, 8:9] <- 1
mat[10, 10] <- 1

diag(mat) <- 2

# The double bridges:

mat[5, 4] <- 1
mat[4, 5] <- 1

mat[7, 8] <- 1
mat[8, 7] <- 1

mat[9, 10] <- 1
mat[10, 9] <- 1

names(mat) <- names(model_temporary$data$fleet.model$gear_group_rel_power)
rownames(mat) <- rownames(model_temporary$data$fleet.model$gear_group_rel_power)

model_temporary$data$fleet.model$gear_group_rel_power <- mat

M <- find_matrix_M(model_temporary)
G <- build_graph(M)

valid <- valid_edge_pairs(G)
ends <- lapply(valid, function(x) ends(G, x))

print(ends)

```

<br><br>

### Bridges for k edges

For k edges,
**Improve Description**

<br><br>

```{r}

valid_edge_cuts <- function(g, k, min_size = 2) {

  edge_ids <- seq_len(ecount(g))

  if (length(edge_ids) < k) {
    return(list())
  }

  edge_sets <- combn(edge_ids, k, simplify = FALSE)

  valid <- list()

  for (edges in edge_sets) {

    g2 <- delete_edges(g, edges)

    comp <- components(g2)

    if (comp$no == 2 && min(comp$csize) >= min_size) {
      valid[[length(valid) + 1]] <- edges
    }
  }

  valid
}

# Need to remove all edges that appear lower than k

```

<br><br>

An example is shown below:

<br><br>

```{r}

# model_temporary <- model
# 
# mat <- as.data.frame(matrix(0, nrow = 12, ncol = 10))
# 
# mat[1:4, 1:4] <- 1
# mat[5:7, 5:7] <- 1
# mat[8:9, 8:9] <- 1
# mat[10, 10] <- 1
# 
# diag(mat) <- 2
# 
# # The bridges:
# 
# mat[5, 4] <- 1
# mat[4, 5] <- 1
# 
# mat[7, 8] <- 1
# mat[8, 7] <- 1
# 
# mat[9, 10] <- 1
# mat[10, 9] <- 1
# 
# names(mat) <- names(model_temporary$data$fleet.model$gear_group_rel_power)
# rownames(mat) <- rownames(model_temporary$data$fleet.model$gear_group_rel_power)
# 
# model_temporary$data$fleet.model$gear_group_rel_power <- mat
# 
# M <- find_matrix_M(model_temporary)
# G <- build_graph(M)
# 
# valid <- valid_edge_cuts(G, 2)
# ends <- lapply(valid, function(x) ends(G, x))
# 
# print(ends)

```

<br><br>

## Acknowledgements

__
__

